Collective behavior of Janus Particles Suspended in a Viscous Fluid

### 2022-07-25 doc/PRF2022/PRF_main.tex
continuting on Sec II

\subsection{Mobility problem}
A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
The domain $\Omega(t) \subset \mathbb{R}^2$ is the solvent phase and $t$
is time. The particles are disks of radius $c$, center $\aa_i$, and
orientation $\theta_i$ relative to the horizontal axis. The boundary of
$\Omega$ is $\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where
$\Gamma_i$ is the boundary of Janus particle $i$. Assuming inertial
terms are negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
viscosity. The solvent velocity satisfies the no-slip boundary condition
for a rigid body motion
\begin{align}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$.
Without inertia, the system in~\eqref{eq:stokes}-\eqref{eq:rigid-vel}
also satisfies the force-free and torque-free conditions.
%This means
%that the total force $\FF_i,$ and torque $T_i$ from the hydrodynamic
%stress on each particle (with subscript `$i$') balance the force and
%torque from the hydrophobic potential~\cite{Fu2022_JFM}. 
The total force
and torque from the fluid flow are
\begin{align}
  \label{eq:force}
 \FF_i = \int_{\Gamma_i} \ssigma \cdot \nnu_i \, \dif s, \quad
 T_i = \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu_i) \, \dif s,
  \quad i=1,\ldots,N_b,
\end{align}
respectively, where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu +
\nabla \uu^T \right)$ is the hydrodynamic stress tensor (pressure
tensor), and $\nnu_i$ is the unit outward normal on the $i$th particle.

@@

The governing equation for $u(\xx,t)$ is obtained by computing the
domain variation of the free energy $F$~\cite{Fu20}. We obtain a
screened Laplace equation boundary value problem as a model for the
hydrophobic interaction between water and JPs:
\begin{alignat}{3}
  \label{eq:SL}
  -\rho^2 \Delta u + u &= 0,\qquad &&\xx \in \Omega, \\
  u &= g, &&\xx \in \bd\Omega, \\
  u &\rightarrow 0, &&|\xx| \rightarrow \infty,
\end{alignat}
where the boundary condition $g$ gives the degree of hydrophobicity the
particle-solvent interface.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary integral represenation and time stepping}
We recast \eqref{eq:stokes}--\eqref{eq:rigid-vel} and \eqref{eq:SL}
as boundary integral equations (BIEs) and discretize each BIE at
$N$ points on each of the $N_b$ particles with a collocation method.
To express the solution of \eqref{eq:SL},
we adopt the double layer potential
\begin{align}
\label{eq:SL_BIE}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy
, \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental solution to the screened Laplace equation \eqref{eq:SL},
$K_0$ is the zeroth-order modified Bessel function of the first kind,
and $\sigma$ is a scalar-valued density function.
The subscript in $\dif s_\yy$ denotes integration with respect to $\yy \in \Gamma$. 
For the velocity, we use the representation [REF]
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$,
Stokeslet 
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr \otimes \rr}{|\rr|^2}\right)$,
and Rotlet
$\RR(\xx,\aa_i) = \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$,
$\rr = \xx - \aa_i$ [REF].

@@

First, we must simplify the
form of the free energy \eqref{eq:free_energy}
using integration by parts and \eqref{eq:SL}
to obtain
\begin{equation}
\label{eq:free_energy2}
F = -\gamma
\int_{\Gamma} \rho g \nabla u \cdot \nnu \,\dif s
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right).
\end{equation}
%
Here, we have substituted $g$ for $u$ since the boundary values are given.
However, evaluating $\nabla u \cdot \nnu$ on $\Gamma$ based on \eqref{eq:SL_BIE}
involves calculating a quadruple layer potential which has a
well-known obstacle in numerical implementation.
To overcome this obstacle, we use
%
\begin{equation}
\label{eq:normal_deriv}
\nabla u(\xx) \cdot \nnu(\xx)=
%\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) =
-\frac{1}{\rho^2} {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx)
+ \frac{\dif}{\dif s}\SSS\left[\frac{\dif \sigma}{\dif s}\right](\xx), \quad \xx \in \Gamma.
\end{equation}
%
Here, ${\bf t}(\xx)$ is the tangent vector and $\dif/\dif s$ is the arclength derivative.
See Appendix in \S\ref{sec:appendix} for the proof and details.
Substituting \eqref{eq:normal_deriv} into \eqref{eq:free_energy2} for the normal derivative
leads to single layer integral which is more straightforward to evaluate and we compute the
arclength derivative using spectrally accurate Fourier transform.
We have validated the form of \eqref{eq:normal_deriv} by extrapolating 
the primal energy \eqref{eq:free_energy} along test curves placed slightly
outside the particles.

>>>>

\subsection{Mobility problem}
A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
The domain $\Omega(t) \subset \mathbb{R}^2$ is the solvent phase and $t$
is time. The particles are disks of radius $c$, center $\aa_i$, and
orientation $\theta_i$ relative to the horizontal axis. The boundary of
$\Omega$ is $\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where
$\Gamma_i$ is the boundary of Janus particle $i$. Assuming inertial
terms are negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
viscosity. The solvent velocity satisfies the no-slip boundary condition
for a rigid body motion
\begin{align}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$.
Without inertia, the system in~\eqref{eq:stokes}--\eqref{eq:rigid-vel}
also satisfies the force-free and torque-free conditions.
%This means
%that the total force $\FF_i,$ and torque $T_i$ from the hydrodynamic
%stress on each particle (with subscript `$i$') balance the force and
%torque from the hydrophobic potential~\cite{Fu2022_JFM}. 
The total force
and torque from the fluid flow are
\begin{align}
  \label{eq:force}
 \FF_i = \int_{\Gamma_i} \ssigma \cdot \nnu_i \, \dif s, \quad
 T_i = \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu_i) \, \dif s,
  \quad i=1,\ldots,N_b,
\end{align}
respectively, where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu +
\nabla \uu^T \right)$ is the hydrodynamic stress tensor (pressure
tensor), and $\nnu_i$ is the unit outward normal on the $i$th particle.

@@

The governing equation for $u(\xx,t)$ is obtained by computing the
domain variation of the free energy $F$~\cite{Fu20}. We obtain a
screened Laplace equation boundary value problem as a model for the
hydrophobic interaction between water and JPs:
\begin{alignat}{3}
  \label{eq:SLpde}
  -\rho^2 \Delta u + u &= 0,\qquad &&\xx \in \Omega, \\
  \label{eq:SLbc}
  u &= g, &&\xx \in \bd\Omega, \\
  \label{eq:SLfarfield}
  u &\rightarrow 0, &&|\xx| \rightarrow \infty,
\end{alignat}
where the boundary condition $g$ gives the degree of hydrophobicity the
particle-solvent interface.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary integral represenation and time stepping}
We recast~\eqref{eq:stokes}--\eqref{eq:rigid-vel}
and~\eqref{eq:SLpde}--\eqref{eq:SLfarfield} as boundary integral
equations (BIEs) and discretize each BIE at $N$ points on each of the
$N_b$ particles with a collocation method. To express the solution
of~\eqref{eq:SLpde}--\eqref{eq:SLfarfield}, we adopt the double layer
potential
\begin{align}
\label{eq:SL_BIE}
  u({\xx}) = \DD[\sigma](\xx) = \int_\Gamma 
    \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, 
    \dif s_\yy, \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental
solution to the screened Laplace equation~\eqref{eq:SLpde}, $K_0$ is the
zeroth-order modified Bessel function of the first kind, and $\sigma$ is
a scalar-valued density function.  The subscript in $\dif s_\yy$ denotes
integration with respect to $\yy \in \Gamma$.  For the velocity, we use
the representation~\cite{pow-mir1987}
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$, Stokeslet
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr
\otimes \rr}{|\rr|^2}\right)$, and Rotlet $\RR(\xx,\aa_i) =
\frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$, $\rr = \xx - \aa_i$.

@@

First, we must simplify the form of the free energy
\eqref{eq:free_energy} using integration by parts and \eqref{eq:SLpde}
to obtain
\begin{equation}
\label{eq:free_energy2}
F = -\gamma
\int_{\Gamma} \rho g \nabla u \cdot \nnu \,\dif s
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right).
\end{equation}
%
Here, we have substituted $g$ for $u$ since the boundary values are given.
However, evaluating $\nabla u \cdot \nnu$ on $\Gamma$ based on \eqref{eq:SL_BIE}
involves calculating a quadruple layer potential which has a
well-known obstacle in numerical implementation.
To overcome this obstacle, we use
%
\begin{equation}
\label{eq:normal_deriv}
\nabla u(\xx) \cdot \nnu(\xx)=
%\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) =
-\frac{1}{\rho^2} {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx)
+ \frac{\dif}{\dif s}\SSS\left[\frac{\dif \sigma}{\dif s}\right](\xx), \quad \xx \in \Gamma.
\end{equation}
%
Here, ${\bf t}(\xx)$ is the tangent vector and $\dif/\dif s$ is the arclength derivative.
See Appendix in \S\ref{sec:appendix} for the proof and details.
Substituting \eqref{eq:normal_deriv} into \eqref{eq:free_energy2} for the normal derivative
leads to single layer integral which is more straightforward to evaluate and we compute the
arclength derivative using spectrally accurate Fourier transform.
We have validated the form of \eqref{eq:normal_deriv} by extrapolating 
the primal energy \eqref{eq:free_energy} along test curves placed slightly
outside the particles.
